%%*****************************************************************************
%% Copyright (c) 2008 Gerd Neugebauer
%%
%% Permission is granted to copy, distribute and/or modify this document
%% under the terms of the GNU Free Documentation License, Version 1.2
%% or any later version published by the Free Software Foundation;
%% with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts.
%%
%%*****************************************************************************
%% $Id:configuration.tex 7067 2008-05-18 11:06:56Z gene $
%%*****************************************************************************
%% Author: Gerd Neugebauer
%%-----------------------------------------------------------------------------


\section{Configuring \ExBib}

\subsection{Locating the Configuration}

If you are looking for the configurations in the distribution or want
to provide a configuration of your own you need to know how a
configuraion is found. \ExBib\ has a general mechanism for finding
resources. Unfortunately this mechanism can not be used for the
configuration since it is defined in the configuration itself. Thus
the location of a configuration uses a simpler search.

There are two places where \ExBib\ looks for a configuration. Those
are mainly located in the \ExBib\ installation directory (cf.
section~\ref{sec:inst.dir}).
\begin{itemize}
\item The current directory may contain a configuration with the
  proper name and path.
\item The directory \File{config} in the \ExBib\ installation
  directory contains an unpacked directory tree. Any configuration in
  this directory tree with a proper name and path is considered.
\item The directory \File{lib} in the \ExBib\ installation directory
  contains jar archives. Any configuration in a jar archive with a
  proper name and path is considered.
\end{itemize}

The top level configuration file is located with the path
\texttt{exbib}. The prefix \texttt{config/} is tried first; then the
name withpout prefix. The configuration is used with the additional
extension \texttt{.xml} and as given. Thus the configuration
\texttt{cfg} passed in as property or via the command line is sought
with the following full names:
\begin{verbatim}
  config/exbib/cfg.xml
  config/exbib/cfg
  exbib/cfg.xml
  exbib/cfg
\end{verbatim}


\subsection{Extracting a Configuration from an Archive}

To get started you may want to examine the configurations in a jar
archive. This can be accomplished with the program \Prog{jar} whoich
is part of the installed Java. We assume that the program can be found
on your search path for executables.

First you can list the files as follows:
\begin{lstlisting}[language=sh]
  # jar -tvf exbib.jar
\end{lstlisting}

Then you can extract the files you want from the archive
\begin{lstlisting}[language=sh]
  # jar -xvf exbib.jar config/exbib/exbib.xml
\end{lstlisting}

Note that a directory tree is created containing the extracted files.


\subsection{The Configuration Format}

The configuration files are XML files. The default configuration file
(\File{exbib.xml}) contains the following definitions:

\begin{lstlisting}[language=XML]
<exbib fallback="bbl">
  <Resource src="path/exbibFinder"/>
  <Processor base="exbib/processor/" default="exbib"/>
  <BblWriter class="org.extex.exbib.core.io.bblio.BblWriter">
    <linelength>79</linelength>
    <indent>  </indent>
  </BblWriter>
  <DB class="org.extex.exbib.core.db.impl.DBImpl">
    <minCrossrefs>2</minCrossrefs>
  </DB>
  <Sorter base="exbib/sorter/" default="locale"/>
  <AuxReader class="org.extex.exbib.core.io.auxio.AuxReaderImpl"/>
  <BibReader class="org.extex.exbib.core.io.bibio.BibReaderImpl"/>
  <BibPrinter base="exbib/printer/" default="bib"/>
</exbib>
\end{lstlisting}


\INCOMPLETE


\endinput
%
% Local Variables: 
% mode: latex
% TeX-master: "../exbib-manual"
% End: 
